### 2018-03-20 CO749-GraphColoring/assignments/A4.tex
revision

	$$\defi(G|H) = \defi(G|G') + \defi(G'|H)$$

@@

	$$v(G|S) - 1 = v(G|H) = v(G|G') + v(G'|H).$$

@@

	$$\defi(G|H) = \defi(G|S) |N(v)\cap V(S)| -3 = \defi(G|S) $$

@@

	&= \frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|H) + (c+1)\defi(G'|H)}{9}\\
	&=\frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|S) + (c+1)\defi(G'|H)}{9}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +\defi(G'|H) + c - 1}{9} \\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +v(G'|H)/c + 1 + c - 1}{9}\\

@@

	$G$ has at least two vertices not in $S$ or at least $3$ edges not in $S$ (otherwise, it is easy to extend any $3$-coloring of $S$ to $G$). Let $e \in \delta(S)$. Then $S+e$ is a strict subgraph of $G$. Let $\phi$ be a $3$-coloring of $S$ which extends to $S+e$ but not to $G$. Let $H$ be the smallest subgraph of $G$ containing $S$ such that $\phi$ does not extend to $H$. So $H$ contains $S+e$, i.e. $H$ contains $S$ as a strict subgraph.

@@

	The more complex case is when $H=G$. In this case we see $G$ is $\phi$-critical as the minimality of our choice of $H$ implies every strict subgraph of $G$ containing $S$ is colorable with an extension of $\phi$. Let $G_S[\phi]$ denote the graph obtained from $G$ by identifying vertices of $S$ according to their color class with respect to $\phi$. Let $x_1, x_2, x_3$ be vertices of $G_S[\phi]$ corresponding to the respective color classes of $\phi$. If necessary we add edges between each $x_i, x_j$ so they form a $K_3$.

@@

	Now we claim $G_S[\phi]$ is not $3$-colorable. Otherwise we can permute a $3$-coloring of $G_S[\phi]$ so it assigns color $i$ to $x_i$. This coloring then corresponds to an extension of $\phi$ which $3$-colors $G$ (by assigning $i$ each vertex of $G$ in $x_i$). Hence $G_S[\phi]$ contains a $4$-critical subgraph $W$. Let $S' = S \cup W - X$, where $X = W \cap \{x_1,x_2,x_3\}$.

@@

	Since $G$ is girth $5$, $T(W) \leq 3$ as the only cycles of length smaller than $5$ in $W$ intersect $\{x_1,x_2,x_3\}$ so there are at most $3$ vertex-disjoint such cycles. Hence

>>>>

	\begin{equation}\label{chain-rule}
	\defi(G|H) = \defi(G|G') + \defi(G'|H)
	\end{equation}

@@

	\begin{equation}\label{vgs-eq}
		v(G|S) - 1 = v(G|H) = v(G|G') + v(G'|H).
	\end{equation}

@@

	\begin{equation}\label{defi-eq}
	\defi(G|H) = \defi(G|S) |N(v)\cap V(S)| -3 = \defi(G|S) 
	\end{equation}

@@

	&= \frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|H) + (c+1)\defi(G'|H)}{9} &\text{by (\ref{chain-rule}) and (\ref{vgs-eq})}\\
	&=\frac{v(G|S) -1 -v(G'|H) -(c+1)\defi(G|S) + (c+1)\defi(G'|H)}{9} &\text{ by (\ref{defi-eq})}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +\defi(G'|H) + c - 1}{9} &\text{ by Theorem \ref{th:def}}\\
	&\geq \frac{v(G|S) -(c+1)\defi(G|S) +v(G'|H)/c + 1 + c - 1}{9} &\text{ by Theorem \ref{th:def}}\\

@@

	$G$ has at least two vertices not in $S$ or at least $3$ edges not in $S$ (otherwise, it is easy to extend any $3$-coloring of $S$ to $G$). Let $\phi$ be a $3$-coloring of $S$ which extends to a strict supergraph of $S$ but not to $G$. Let $H$ be the smallest subgraph of $G$ containing $S$ such that $\phi$ does not extend to $H$. So $H$ contains $S$ as a strict subgraph.

@@

	The more complex case is when $H=G$. In this case we see $G$ is $\phi$-critical as the minimality of our choice of $H$ implies every strict subgraph of $G$ containing $S$ is colorable with an extension of $\phi$. Let $G_S[\phi]$ denote the graph obtained from $G$ by identifying vertices of $S$ according to their color class with respect to $\phi$. Let $x_1, x_2, x_k$, $k\leq 3$, be vertices of $G_S[\phi]$ corresponding to the respective color classes of $\phi(S)$. If necessary we add edges between each $x_i, x_j$ so they form a complete subgraph.

@@

	Now we claim $G_S[\phi]$ is not $3$-colorable. Otherwise we can permute a $3$-coloring of $G_S[\phi]$ so it assigns color $i$ to each $x_i$. This coloring then corresponds to an extension of $\phi$ which $3$-colors $G$ (by assigning $i$ each vertex of $G$ in $x_i$). Hence $G_S[\phi]$ contains a $4$-critical subgraph $W$. Let $S' = S \cup W - X$, where $X = W \cap \{x_1, \dots, x_k\}$.

@@

	Since $G$ is girth $5$, $T(W) \leq 3$ as the only cycles of length smaller than $5$ in $W$ intersect $\{x_1,\dots, x_k\}$ so there are at most $3$ vertex-disjoint such cycles. Hence
